---
title: "Tidy the output from density estimation models and ECOMOD"
author: "Alexa Varah"
date: "2025-03-31"
output: html_document
---

# 1. Load packages
```{r load-packages, message=FALSE}
rm(list=ls())
library(tidyverse) # for data manipulation; includes dplyr, ggplot2, tidyr, stringr & more
```

# 2. Load & tidy density data
Tidy the simulated density data for both the original model runs and for the 'worst case' model runs.  
In the original density simulations, density estimation models were initialised with 'current' black-grass densities; in other words, they were initialised with the initial density distribution for which each strategy was designed.  
In the 'worst case' density simulations, density estimation models were initialised with very high density distributions for all strategies.  
  
Both data frames need columns for (a) initial density and resistance, (b) soil type, and (c) region, that match those in the economic data.  
  
For the 'worst case' scenario, all simulations were run once with initial high density and then again with initial very high density. **Only the very high density simulations are used in this analysis.**  
The initial density distributions for each state are as follows:   
* "very high"  A 0.1, L 0.1, M 0.2, H 0.3, VH 0.3  
* "high"       A 0.2, L 0.2, M 0.2, H 0.2, VH 0.2  
  
The 'worst case' simulated density data is an R list, nested by initial density and the sequential strategy:   
$VD_100 = all strategies at Very high density (we use only this for the worst case analysis)  
$HD_100 = all strategies at High density (we won't use this in the end as it's not the 'worst' case)  
  
elements 1-4 e.g. $VD_100[[1:4]] are then the four sequential strategies, as follows:  
MIT - MIT - MIT  
BAU - BAU - BAU  
BAU - BAU - MIT  
BAU - MIT - MIT  
  
where each rotation is 6 years (making 18 in total).  
  
As before, there are 37 iterations of each strategy.  
  
## 2.1. Load density data
```{r load-density-data}
density_data <- readRDS("../data/04012024_simulation_results.rds") # original
#d2_worstcase <- readRDS("../data/2023-12-19_sequential_rotation_results.rds") # worst case

#purrr::map(d2_worstcase, tibble::as_tibble, .name_repair = "minimal")

# HD_MMM <- as.data.frame(d2_worstcase$HD_100[[1]]) %>% 
#   dplyr::mutate(Robden_seq = "HD_MMM")
# HD_BBB <- as.data.frame(d2_worstcase$HD_100[[2]]) %>% 
#   dplyr::mutate(Robden_seq = "HD_BBB")
# HD_BBM <- as.data.frame(d2_worstcase$HD_100[[3]]) %>% 
#   dplyr::mutate(Robden_seq = "HD_BBM")
# HD_BMM <- as.data.frame(d2_worstcase$HD_100[[4]]) %>% 
#   dplyr::mutate(Robden_seq = "HD_BMM")
# 
# VD_MMM <- as.data.frame(d2_worstcase$VD_100[[1]]) %>% 
#   dplyr::mutate(Robden_seq = "VD_MMM")
# VD_BBB <- as.data.frame(d2_worstcase$VD_100[[2]]) %>% 
#   dplyr::mutate(Robden_seq = "VD_BBB")
# VD_BBM <- as.data.frame(d2_worstcase$VD_100[[3]]) %>% 
#   dplyr::mutate(Robden_seq = "VD_BBM")
# VD_BMM <- as.data.frame(d2_worstcase$VD_100[[4]]) %>% 
#   dplyr::mutate(Robden_seq = "VD_BMM")

# Join
#simden_wc <- rbind(HD_MMM, HD_BBB, HD_BBM, HD_BMM, VD_MMM, VD_BBB, VD_BBM, VD_BMM)

# Clean up
#rm(HD_MMM, HD_BBB, HD_BBM, HD_BMM, VD_MMM, VD_BBB, VD_BBM, VD_BMM, d2_worstcase)
```

## 2.2. Tidy density data  
For worst case scenario we only want simulations with initial very high density.  
Need to create columns for initial den-res state, soil type, region, sequence of rotations, and, for the worst case data, which rotation (BAU or MIT) was run in the first 6 years of the 18 year sequence.
```{r tidy-density-data}
simden <- density_data |>  
            dplyr::rename(year = iter) |> 
            dplyr::mutate(initdenres = str_extract(strategy , "VD_HR|HD_HR|MD_HR|MD_LR|LD_HR|LD_LR"),
                          initdenres = stringr::str_replace_all(initdenres, "_","-"),
                          initdenres = factor(initdenres , # specify order of levels
                                              levels = c("VD-HR", "HD-HR", "MD-HR", 
                                                         "MD-LR", "LD-HR", "LD-LR")) , 
                          soiltype = str_extract(strategy , "heavy|medium|light"),
                          soiltype = dplyr::recode(soiltype,  # rename levels of 'soiltype'
                                                   "heavy" = "h", #old, new
                                                   "medium" = "m", #old, new
                                                   "light" = "l"), #old, new
                          soiltype = factor(soiltype, # specify order of levels
                                          levels = c("h", "m", "l")),
                          region = str_extract(strategy , "north|central|east"),
                          region = dplyr::recode(region, # rename levels of 'region'
                                                 "north" = "n", #old, new
                                                 "central" = "c", #old, new
                                                 "east" = "e"), #old, new
                          region = factor(region, # specify order of levels
                                          levels = c("n", "c", "e")),
                          strategy = stringr::str_replace_all(strategy, "FF","MIT"),
                          scenario = str_extract(strategy , "BAU|CWW|MIT"),
                          year = as.numeric(year),
                          # Create new ID on which to later join this data frame with the economic data
                          ID = paste(scenario, region, soiltype, initdenres, year, sep="_")) |> 
            ungroup() |>  
            # Sum the proportions of absent and low density black-grass
            rowwise() |> 
            dplyr::mutate(al = sum(c_across(a:l), na.rm = T)) %>% # new category: absent & low
            ungroup() |>
            dplyr::relocate(al, .before=m) |>
            dplyr::select(-c(step)) |>
            dplyr::relocate(scenario, .after = strategy) |>
            dplyr::relocate(soiltype, .after = region)
            

# Clean up
rm(density_data)
```


# 3. Load & tidy ECOMOD data

## 3.1. MIT
```{r load-and-prep-MIT}
## 1. MIT - Create long-format data frames.

## MIT - Low density model run
mit_low <- read.csv("../output/mit_ECOMOD_output_L_combined.csv", header=TRUE) %>% 
  dplyr::select(-X) %>% # remove extra column added by R when reading in the file
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

mit_low <- reshape(mit_low, varying=c(2:85), direction="long", 
                  idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "low") # add a column to indicate the model run


## MIT - Medium density model run

mit_med <- read.csv("../output/mit_ECOMOD_output_M_combined.csv", header=TRUE) %>% 
  dplyr::select(-X) %>% # remove extra column added by R when reading in the file
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

mit_med <- reshape(mit_med, varying=c(2:85), direction="long", 
                  idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "medium") # add a column to indicate the model run


## MIT - High density model run

mit_high <- read.csv("../output/mit_ECOMOD_output_H_combined.csv", header=TRUE) %>% 
  dplyr::select(-X) %>% # remove extra column added by R when reading in the file
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

mit_high <- reshape(mit_high, varying=c(2:85), direction="long", 
                   idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "high") # add a column to indicate the model run


## MIT - Very high density model run

mit_vhig <- read.csv("../output/mit_ECOMOD_output_V_combined.csv", header=TRUE) %>% 
  dplyr::select(-X) %>% # remove extra column added by R when reading in the file
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

mit_vhig <- reshape(mit_vhig, varying=c(2:85), direction="long", 
                   idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "veryhigh") # add a column to indicate the model run


## 2. MIT - Join data frames.

# Need to create initcondit col in the same format as in the sim data.
# Also need to add a column that is identical in both the economic and simulation data frames, 
# and which identifies scenario, initial conditions and year.

mit_econdata <- rbind(mit_low, mit_med, mit_high, mit_vhig) %>% 
  dplyr::mutate(
    field_name = gsub('\\.', '_', field_name), # replace . with _
    initdenres = sub('_.*', '', field_name), # extract chrs before first '_'
    soiltype = sub('.*\\_', '', field_name), # extract chrs after last '_'
    region = substr(field_name, start = 7, stop = 7), # extract 7th character
    region = tolower(region),
    initcondit = paste(region, soiltype, initdenres, sep = "_"),
    scenario = "MIT",
    ID = paste(scenario, initcondit, rotation, sep = "_")
    )

rm(mit_med, mit_high, mit_vhig)
```

## 3.2. BAU
```{r load-and-prep-BAU}
## 1. BAU - Create long-format data frames.

## BAU - Low density model run

bau_low <- read.csv("../output/bau_ECOMOD_output_L.csv", header=TRUE) %>% 
  dplyr::select(-Field_no) %>% # remove unnecessary column
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

bau_low <- reshape(bau_low, varying=c(2:85), direction="long", 
                  idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "low") # add a column to indicate the model run


## BAU - Medium density model run

bau_med <- read.csv("../output/bau_ECOMOD_output_M.csv", header=TRUE) %>% 
  dplyr::select(-Field_no) %>% # remove unnecessary column
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

bau_med <- reshape(bau_med, varying=c(2:85), direction="long", 
                  idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "medium") # add a column to indicate the model run


## BAU - High density model run

bau_high <- read.csv("../output/bau_ECOMOD_output_H.csv", header=TRUE) %>% 
  dplyr::select(-Field_no) %>% # remove unnecessary column
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

bau_high <- reshape(bau_high, varying=c(2:85), direction="long", 
                   idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "high") # add a column to indicate the model run


## BAU - Very high density model run

bau_vhig <- read.csv("../output/bau_ECOMOD_output_V.csv", header=TRUE) %>% 
  dplyr::select(-Field_no) %>% # remove unnecessary column
  rename_with(stringr::str_replace, # add '.' before the number in 'crop' colnames
              pattern = "crop", replacement = "crop.", 
              matches("crop")) %>% 
  rename(meanrotgrossprof = totalrotgrossprof) %>% 
  data.frame()

bau_vhig <- reshape(bau_vhig, varying=c(2:85), direction="long", 
                   idvar = "field_name", sep = ".", timevar = "rotation") %>% 
  arrange(field_name) %>% 
  mutate(modelrun = "veryhigh") # add a column to indicate the model run


## 2. BAU - Join data frames.

# Need to create initcondit col in the same format as in the density data.
# Also need to add a column that is identical in both the economic and simulation data frames, 
# and which identifies scenario, initial conditions and year.

bau_econdata <- rbind(bau_low, bau_med, bau_high, bau_vhig) %>% 
  dplyr::mutate(
    initdenres = sub('.*\\_', '', field_name), # extract chrs after last '_'
    soiltype = sapply(strsplit(field_name, "_"), function(x) x[2]), # extract characters between 1st and 2nd underscore ('x[2]' retrieves the second result)
    soiltype = substr(soiltype, start = 1, stop = 1), # extract 1st character
    region = sub('_.*', '', field_name), # extract chrs before first '_'
    region = tolower(region),
    initcondit = paste(region, soiltype, initdenres, sep = "_"),
    scenario = "BAU",
    ID = paste(scenario, initcondit, rotation, sep = "_")
  )

# Clear up
rm(bau_med, bau_high, bau_vhig)
```


## 3.3. Join economic scenarios & tidy
```{r join-two-economic-scenarios}
econdata <- rbind(mit_econdata, bau_econdata) %>% 
  dplyr::select(-field_name)

## Clean up
rm(bau_econdata)
rm(mit_econdata)
```


# 4. Join datasets and calculate weighted values (current densities)
When joining, the `econdata` rows will automatically repeat to fill the 37 imputations present in the density data.  

Only winter wheat yield will vary depending on black-grass density; yield of all other crops will be unaffected by the weighting process. I calculate weighted yield, output, fuel cost, labour cost and operations cost. Fuel, labour and operations costs (in ECOMOD, Operation cost = Fuel cost + Labour cost) all involve using work rates in their calculation. Work rate is a function of crop yield and so these variables  will be affected by black-grass density when the crop is wheat. When you do the join, the `econdata` rows will automatically repeat to fill the 37 iterations present in `simdata`.  

Not all variables need to be weighted as they are not all affected by crop yield and will therefore not respond to changes in BG density. Furthermore, even weighted variables won't always show differences between imputations as at low and medium densities, black-grass doesn't affect yield.  
  
Opportunity cost is calculated as BBB-BBM, BBB-BMM, or BBB-MMM. This means a negative value indicates that BBB (i.e. 3 consecutive BAU rotations) had higher gross profit, so the other sequence (BBM, BMM or MMM) makes the farmer less money.

## 4.1. Calculate weighted values (including opportunity and productivity costs)  
Calculate weighted gross profit, all other weighted variables, opportunity cost and productivity costs.  
```{r calculate-weighted-values-and-oc}
wtd_values_pre <- left_join(simden %>% dplyr::filter(scenario != "CWW"), 
                            econdata, 
                            by="ID", 
                            relationship = "many-to-many") %>% 
  
  # Calculate weighted gross profit for each row for the affected variables
  dplyr::mutate(
    wtd_gp = case_when(
      modelrun == "low" ~ al*grossprof,
      modelrun == "medium" ~ m*grossprof,
      modelrun == "high" ~ h*grossprof,
      modelrun == "veryhigh" ~ v*grossprof
    ),
    wtd_fuelcost = case_when(
      modelrun == "low" ~ al*fuelcost,
      modelrun == "medium" ~ m*fuelcost,
      modelrun == "high" ~ h*fuelcost,
      modelrun == "veryhigh" ~ v*fuelcost
    ),
    wtd_labcost = case_when(
      modelrun == "low" ~ al*labcost,
      modelrun == "medium" ~ m*labcost,
      modelrun == "high" ~ h*labcost,
      modelrun == "veryhigh" ~ v*labcost
    ),
    wtd_opscost = case_when(
      modelrun == "low" ~ al*opcost,
      modelrun == "medium" ~ m*opcost,
      modelrun == "high" ~ h*opcost,
      modelrun == "veryhigh" ~ v*opcost
    ),
    wtd_yield = case_when(
      modelrun == "low" ~ al*yield,
      modelrun == "medium" ~ m*yield,
      modelrun == "high" ~ h*yield,
      modelrun == "veryhigh" ~ v*yield
    ),
    wtd_output = case_when(
      modelrun == "low" ~ al*output,
      modelrun == "medium" ~ m*output,
      modelrun == "high" ~ h*output,
      modelrun == "veryhigh" ~ v*output
    ),
    ID = factor(ID)
  ) %>%
  
  # Add weighted variables across model runs (low, med, high & vhigh den) 
  # to get total gross profit, yield etc (i.e., add the contributions from the 
  # different densities in different parts of the field) for each year of each 
  # strategy (ID), and for each imputation of that year-strategy (imp).
  dplyr::group_by(imp, ID) %>% 
  dplyr::mutate(
    wtd_gp = sum(wtd_gp),
    wtd_fuelcost = sum(wtd_fuelcost),
    wtd_labcost = sum(wtd_labcost),
    wtd_opscost = sum(wtd_opscost),
    wtd_yield = sum(wtd_yield),
    wtd_output = sum(wtd_output)
  ) %>% 
  ungroup() %>% 
  distinct(imp, year, strategy, .keep_all = T) %>% 
  dplyr::select(
    imp, strategy, scenario.x, year, 
    region.x, soiltype.x, initdenres.x, initcondit, ID,
    crop, wtd_gp:wtd_output
  ) %>% 
  dplyr::rename_with(~ stringr::str_remove(., ".x"), everything()) %>% # remove .x in colnames
  
  # Specify formats and order of levels (so later plots are in correct order):
  dplyr::mutate( 
    scenario = factor(scenario, levels=c("BAU", "MIT")),
    initdenres = factor(initdenres, levels = c("VD-HR", "HD-HR", "MD-HR",
                                               "MD-LR", "LD-HR", "LD-LR")),
    initcondit = factor(initcondit, levels = c(
      "n_h_VD-HR", "n_m_VD-HR", "n_l_VD-HR", 
      "c_h_VD-HR", "c_m_VD-HR", "c_l_VD-HR",
      "e_h_VD-HR", "e_m_VD-HR", "e_l_VD-HR",
      
      "n_h_HD-HR", "n_m_HD-HR", "n_l_HD-HR",
      "c_h_HD-HR", "c_m_HD-HR", "c_l_HD-HR",
      "e_h_HD-HR", "e_m_HD-HR", "e_l_HD-HR",
      
      "n_h_MD-HR", "n_m_MD-HR", "n_l_MD-HR",
      "c_h_MD-HR", "c_m_MD-HR", "c_l_MD-HR",
      "e_h_MD-HR", "e_m_MD-HR", "e_l_MD-HR",
      
      "n_h_MD-LR", "n_m_MD-LR", "n_l_MD-LR",
      "c_h_MD-LR", "c_m_MD-LR", "c_l_MD-LR",
      "e_h_MD-LR", "e_m_MD-LR", "e_l_MD-LR",
      
      "n_h_LD-HR", "n_m_LD-HR", "n_l_LD-HR",
      "c_h_LD-HR", "c_m_LD-HR", "c_l_LD-HR",
      "e_h_LD-HR", "e_m_LD-HR", "e_l_LD-HR",
      
      "n_h_LD-LR", "n_m_LD-LR", "n_l_LD-LR",
      "c_h_LD-LR", "c_m_LD-LR", "c_l_LD-LR",
      "e_h_LD-LR", "e_m_LD-LR", "e_l_LD-LR"))
  ) %>% 
  ungroup() %>% 
  
  # Calculate opportunity cost of switching out of BAU
  group_by(imp, year, initcondit) %>% 
  dplyr::mutate(wtd_opp_cost = wtd_gp - wtd_gp[scenario == "BAU"]) %>% 
  ungroup() %>% 
  arrange(scenario, initcondit, year) %>% 
  
  # Calculate productivity cost of switching out of BAU
  dplyr::mutate(ww_yield = ifelse(stringr::str_detect(crop,'wheat'),wtd_yield,0)) %>% # for any crop that isn't wheat, assign zero yield
  group_by(imp, year, initcondit) %>% 
  dplyr::mutate(wtd_prod_cost = ww_yield - ww_yield[scenario == "BAU"]) %>% 
  ungroup() %>% 
  arrange(scenario, initdenres, region, soiltype, year)  


# Join other costs (that didn't need to be weighted) back on
econdata_small <- econdata %>% 
  dplyr::select(ID, fertcost, seedcost, herbcost, sundry, varcost) %>% 
  distinct(ID, .keep_all=TRUE)

wtd_values <- inner_join(econdata_small, wtd_values_pre,  by = "ID") %>% 
  dplyr::relocate(
    imp, ID, scenario, region, soiltype, initdenres, initcondit, year, crop,
    wtd_gp, wtd_opp_cost, wtd_yield, ww_yield, wtd_prod_cost, wtd_output, wtd_fuelcost, wtd_labcost, wtd_opscost) %>% 
  arrange(scenario, initdenres, region, soiltype) %>% 
  # Create variable for 'other' costs (to be used in a plot later):
  dplyr::mutate(other = sundry + fertcost + seedcost) %>% 
  dplyr::relocate(other, .after = varcost)

# Clean up
rm(econdata_small, wtd_values_pre)
```

## 4.2. Calculate cost due to lost wheat yield (original)
  
The yield from low density model runs is essentially the 'potential' output as low black-grass densities don't impact wheat yield.  
Costs due to lost wheat yield are therefore:  
potential output (no BG) - output (BG present)  
```{r calculate-costs-due-to-yield-loss}
# Process and join low density model runs
mit_low_tidy <- mit_low %>% 
  dplyr::mutate(
    year = rotation,
    field_name = gsub('\\.', '_', field_name), # replace . with _
    initdenres = sub('_.*', '', field_name), # extract chrs before first '_'
    soiltype = sub('.*\\_', '', field_name), # extract chrs after last '_'
    region = substr(field_name, start = 7, stop = 7), # extract 7th character
    region = tolower(region),
    initcondit = paste(region, soiltype, initdenres, sep = "_"),
    scenario = "MIT",
    ID = paste(scenario, initcondit, rotation, sep = "_")
    ) %>% 
  dplyr::select(-c(meanrotgrossprof, field_name, rotation))

bau_low_tidy <- bau_low %>% 
  dplyr::mutate(
    year = rotation,
    initdenres = sub('.*\\_', '', field_name), # extract chrs after last '_'
    soiltype = sapply(strsplit(field_name, "_"), function(x) x[2]), # extract string between 1st and 2nd underscore ('x[2]' pulls out the 2nd split string)
    soiltype = substr(soiltype, start = 1, stop = 1), # extract 1st character
    region = sub('_.*', '', field_name), # extract chrs before first '_'
    region = tolower(region),
    initcondit = paste(region, soiltype, initdenres, sep = "_"),
    scenario = "BAU",
    ID = paste(scenario, initcondit, rotation, sep = "_")
  ) %>% 
  dplyr::select(-c(meanrotgrossprof, field_name, rotation))


# Join these
econdata_low <- rbind(mit_low_tidy, bau_low_tidy) 

# Rename output column
econdata_low <- econdata_low %>% dplyr::rename(potential_output = output)

# Calculate cost due to lost yield
# Join output column from econdata_low_tidy onto weighted_orig, then do the subtraction.
wtd_values <- inner_join(wtd_values, econdata_low %>% dplyr::select(ID, potential_output), by = "ID") %>%
  dplyr::mutate(
    wtd_cyl = potential_output - wtd_output,
    wtd_cyl = round(wtd_cyl, digits = 2)
    )

rm(econdata_low, bau_low, bau_low_tidy, mit_low, mit_low_tidy)
```


# 5. Save output files
```{r save-output}
saveRDS(simden, file = "../output/simden_tidy.rds")
saveRDS(wtd_values, file = "../output/weighted-gp-yield-oc.rds")
saveRDS(econdata, file = "../output/econdata.rds")
```
