---
title: "Analyse trade-offs"
author: "Alexa Varah"
date: "2024-05-21"
output: html_document
---

# 1. Intro
  
This script looks at what happens to farm gross profits, yield (in calories), and carbon emissions when farmers switch from a BAU crop rotation to strategies aimed at mitigating herbicide-resistance black-grass in England.  
  
We ignore the 'continuous winter wheat' scenario from the previous paper as we showed there that it's worst for controlling herbicide-resistant black-grass.  

Data `c_gp_yld` is already summarised across imputations.  


## Sources of emissions in CFT

**Pesticide emissions**  
Pesticide emissions = number of pesticide applications * embodied emissions in pesticides (kg COs-eq / dose / ha) * crop yield / ha  
This doesn't account for the fuel used when applying pesticides, only the emissions caused by pesticide production.  

**Direct and indirect field N2O**  
Direct and indirect field N2O = Background direct and indirect N2O + Fertiliser induced field emissions (FIE)  

Background direct and indirect N2O = N2O + NO (they don't include NH3 volatilisation + leaching in this equation)
Fertiliser induced field emissions = N to N2O conversion factor * (N2O + NO + NH3 volatilisation + leaching)
NH3 volatilisation = volatilisation emissions factor of 0.01 * 
                      sum of NH3 volatilisation from each fertiliser application
NH3 volatilisation from a fertiliser application = amount of N applied * 
      EXP(sum of NH3 volatilisation factors for climate, crop type, soil texture, SOM, soil CEC, drainage, pH + 
              NH3 from fertiliser production emissions + 
                NH3 from fertiliser application)
              
**Field energy use**  
This contains emissions from diesel used in tillage, herb spraying, fert spreading/spraying, other field ops (harvest etc - we haven't input data on these other field ops).

  
# 2. Load packages, functions, data
## 2.1. Load packages & functions
```{r}
# Load packages
rm(list = ls(all = TRUE)) # clear envt
#library(hrbrthemes) # for theme_ipsum() in plot
library(tidyverse)
library(patchwork) # for panel plots
library(grateful)
library(readxl) # for reading in Excel workbooks / spreadsheets
library(viridis)
library(forcats)

# Function to count complete observations ----
nobs <- function(x) length(x[!is.na(x)])

# Function to annotate plots (N.B. I no longer use this) ----
# annotate_text2 <- function(label, x, y, facets=NULL, hjust=0, vjust=0, color='black', alpha=NA,
#                           family=thm$text$family, size=thm$text$size, fontface=1, lineheight=1.0,
#                           box_just=ifelse(c(x,y)<0.5,0,1), margin=unit(size/2, 'pt'), thm=theme_get()) {
#   x <- scales::squish_infinite(x)
#   y <- scales::squish_infinite(y)
#   data <- if (is.null(facets)) data.frame(x=NA) else data.frame(x=NA, facets)
# 
#   tg <- grid::textGrob(
#     label, x=0, y=0, hjust=hjust, vjust=vjust,
#     gp=grid::gpar(col=alpha(color, alpha), fontsize=size, fontfamily=family, fontface=fontface, lineheight=lineheight)
#   )
#   ts <- grid::unit.c(grid::grobWidth(tg), grid::grobHeight(tg))
#   vp <- grid::viewport(x=x, y=y, width=ts[1], height=ts[2], just=box_just)
#   tg <- grid::editGrob(tg, x=ts[1]*hjust, y=ts[2]*vjust, vp=vp)
#   inner <- grid::grobTree(tg, vp=grid::viewport(width=unit(1, 'npc')-margin*2, height=unit(1, 'npc')-margin*2))
# 
#   layer(
#     data = NULL,
#     stat = StatIdentity,
#     position = PositionIdentity,
#     geom = GeomCustomAnn,
#     inherit.aes = TRUE,
#     params = list(
#       grob=grid::grobTree(inner), 
#       xmin=-Inf, 
#       xmax=Inf, 
#       ymin=-Inf, 
#       ymax=Inf
#     )
#   )
# }

```

## 2.2. Load data
```{r}
c_gp_yld <- read.csv("../output/trade_offs_tidy.csv", header = T) %>% 
  dplyr::mutate(
    scenario = factor(scenario, levels = c("BAU", "MIT")),
    region = factor(region, levels = c("north", "central", "east")),
    soiltype = factor(soiltype, levels = c("heavy", "medium","light")),
    initDR = factor(initDR, levels = c("LD-LR", "LD-HR", "HD-HR"))
    )
```

# 3. Summary table in paper: Percentage change in gp, C, calories, landuse, for each level of initDR
## 3.1. Calculate % change from rotation means for 3 levels of initDR
Here, I calculate the mean gp, kcal, land use, and GHG emissions across the rotation, then calculate the difference between BAU and MIT for each rotation, then calculate the % change for the rotation. The min and max values show the range across regions and soil types.  
```{r}
pc_change_rotn <- c_gp_yld %>%
  
  # 1. Calculate rotation mean for gp, yield, landuse, GHG emiss
  dplyr::group_by(scenario, region, soiltype, initDR) %>% 
  dplyr::mutate(
    # gross profit mean
    gp_mean = mean(gp),
    cal_mean = mean(calories),
    lu_mean = mean(land_use),
    ghg_mean = mean(c_total)
  ) %>% 
  ungroup() %>% 
  dplyr::select(c(scenario:initDR, gp_mean:ghg_mean)) %>% 
  dplyr::distinct(scenario, region, soiltype, initDR, .keep_all = TRUE) %>% 
  
  # 2. Calculate difference between MIT and BAU
  dplyr::group_by(region, soiltype, initDR) %>% 
  dplyr::mutate(
    gp_diff = gp_mean - gp_mean[scenario == "BAU"],
    cal_diff = cal_mean - cal_mean[scenario == "BAU"],
    lu_diff = lu_mean - lu_mean[scenario == "BAU"],
    ghg_diff = ghg_mean - ghg_mean[scenario == "BAU"]
  ) %>% 
  ungroup() %>% 
  
  # 3. Calculate percentage change
  dplyr::group_by(region, soiltype, initDR) %>% 
  dplyr::mutate(
    
    # gross profit % change
    gp_pc = (gp_diff/gp_mean[scenario == "BAU"])*100, # opportunity cost is already the difference in gross profit 
    
    # calories % change
    kcal_pc = (cal_diff/cal_mean[scenario == "BAU"])*100,    
    
    # land use % change
    landuse_pc = (lu_diff/lu_mean[scenario == "BAU"])*100,  
    
    # carbon emissions % change
    ghg_pc = (ghg_diff/ghg_mean[scenario == "BAU"])*100
    
  ) %>% 
  ungroup() %>% 
  
  # 4. Calculate min and max values (this grouping will give range across region and soil types)
  dplyr::filter(scenario == "MIT") %>% 
  dplyr::group_by(initDR) %>%
  dplyr::mutate(

    # gross profit difference min & max
    gp_pc_min = min(gp_pc, na.rm = TRUE),
    gp_pc_max = max(gp_pc, na.rm = TRUE),

    # calories difference min & max
    kcal_pc_min = min(kcal_pc, na.rm = TRUE),
    kcal_pc_max = max(kcal_pc, na.rm = TRUE),

    # land use difference min & max
    landuse_pc_min = min(landuse_pc, na.rm = TRUE),
    landuse_pc_max = max(landuse_pc, na.rm = TRUE),

    # GHG difference min & max
    ghg_pc_min = min(ghg_pc, na.rm = TRUE),
    ghg_pc_max = max(ghg_pc, na.rm = TRUE)

  ) %>%
  ungroup() %>%
  
  # 5. Calculate mean for each level of initDR
  dplyr::group_by(initDR) %>% 
  dplyr::mutate(
    
    # gross profit rotation mean % change
    gp_pc_mean = mean(gp_pc),

    # calories rotation mean % change
    kcal_pc_mean = mean(kcal_pc),
    
    # land use rotation mean % change
    lu_pc_mean = mean(landuse_pc),
    
    # carbon emissions rotation mean % change
    ghg_pc_mean = mean(ghg_pc)
    
  ) %>% 
  
  #Select just useful columns and filter out duplicates
  #dplyr::select(initDR, contains("pc_")) %>% 
  dplyr::select(initDR, 
                gp_pc_mean, gp_pc_min, gp_pc_max,
                kcal_pc_mean, kcal_pc_min, kcal_pc_max,
                lu_pc_mean, landuse_pc_min, landuse_pc_max,
                ghg_pc_mean, ghg_pc_min, ghg_pc_max) %>% 
  dplyr::distinct(initDR, .keep_all = TRUE) %>% 
  arrange(initDR)

pc_change_rotn
```

This approach (in which I calculated % change from rotation means) is the same as the approach used in the previous paper. This approach also makes more sense than the one below as here, the calories produced per m2 match up better with the landuse. If you produce more calories per m2 in MIT, then you should require less land to produce 1000 kcal. This is reflected here: where the % change is +ve for calories, it's -ve for landuse. Using the first approach, this relationship isn't there.

## 3.2. Save output as .csv
```{r}
write.csv(pc_change_rotn, "../output/summarytable_from_rotn_means.csv", row.names = FALSE)
```



# 4. Mean difference in C emissions per initDR
This gives min and max values across regions and soil types. Used in the paper.  
```{r}
c_gp_yld %>%
  
  # 1. Calculate rotation mean gp, yield, landuse, GHG emiss
  dplyr::group_by(scenario, region, soiltype, initDR) %>% 
  dplyr::mutate(
    ghg_mean = mean(c_total)
  ) %>% 
  ungroup() %>% 
  dplyr::select(c(scenario:initDR, ghg_mean)) %>% 
  dplyr::distinct(scenario, region, soiltype, initDR, .keep_all = TRUE) %>% 
  
  # 2. Calculate difference between MIT and BAU
  dplyr::group_by(region, soiltype, initDR) %>% 
  dplyr::mutate(
    ghg_diff = ghg_mean - ghg_mean[scenario == "BAU"]
  ) %>% 
  ungroup() %>% 
  
  # 3. Calculate min and max values (this grouping will give range across region and soil types)
  dplyr::filter(scenario == "MIT") %>% 
  dplyr::group_by(initDR) %>%
  dplyr::mutate(
    ghg_diff_min = min(ghg_diff, na.rm = TRUE),
    ghg_diff_max = max(ghg_diff, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
  # 5. Calculate mean for each level of initDR
  dplyr::group_by(initDR) %>% 
  dplyr::mutate(
    ghg_diff_mean = mean(ghg_diff)
  ) %>% 
  
  #Select just useful columns and filter out duplicates
  dplyr::select(initDR, 
                ghg_diff_mean, ghg_diff_min, ghg_diff_max) %>% 
  dplyr::distinct(initDR, .keep_all = TRUE) %>% 
  arrange(initDR)
```


# 5. Summarise across rotations 
For plots of the effect of initial density-resistance state *within regions and soil types* (i.e., panel plots, where panels are region and soil), range needs to show variability across years in the rotation.  
  
```{r, attr.source='.numberLines'}
trdoff_rot <- c_gp_yld %>% 
  dplyr::group_by(scenario, region, soiltype, initDR) %>% 
  # calculate min and max values (this grouping will give range across years)----
  dplyr::mutate(
    gp_min = min(gp, na.rm = TRUE),
    gp_max = max(gp, na.rm = TRUE),
    oc_min = min(oc, na.rm = TRUE),
    oc_max = max(oc, na.rm = TRUE),
    ww_pc_min = min(ww_pc, na.rm = TRUE),
    ww_pc_max = max(ww_pc, na.rm = TRUE),
    calories_min = min(calories, na.rm = TRUE),
    calories_max = max(calories, na.rm = TRUE),
    calories_diff_min = min(calories_diff, na.rm = TRUE),
    calories_diff_max = max(calories_diff, na.rm = TRUE),
    yield_min = min(yield, na.rm = TRUE),
    yield_max = max(yield, na.rm = TRUE),
    land_use_min = min(land_use, na.rm = TRUE),
    land_use_max = max(land_use, na.rm = TRUE),
    land_use_diff_min = min(land_use_diff, na.rm = TRUE),
    land_use_diff_max = max(land_use_diff, na.rm = TRUE),
    c_total_min = min(c_total, na.rm = TRUE),
    c_total_max = max(c_total, na.rm = TRUE),
    c_emiss_diff_min = min(c_emiss_diff, na.rm = TRUE),
    c_emiss_diff_max = max(c_emiss_diff, na.rm = TRUE)
  ) %>% 
  # calculate mean values---- 
  dplyr::summarise(across(c(
    gp, gp_min, gp_max, # all min values are the same so the mean == min, ditto for max
    oc, oc_min, oc_max, 
    ww_pc, ww_pc_min, ww_pc_max, 
    calories, calories_min, calories_max,
    calories_diff, calories_diff_min, calories_diff_max,
    yield, yield_min, yield_max,
    land_use, land_use_min, land_use_max,
    land_use_diff, land_use_diff_min, land_use_diff_max,
    c_total, c_total_min, c_total_max,
    c_emiss_diff, c_emiss_diff_min, c_emiss_diff_max
    ),.fns = list(
      mean=~mean(.x, na.rm = TRUE),
      n=~nobs(.x)))) %>% 
  ungroup() %>%
  # tidy up----
  dplyr::select(-contains(c('min_n', 'max_n'))) %>% 
  dplyr::rename_at(vars(everything()), ~ sub("min_mean", "min", .x)) %>% # remove '_mean' in 'minimum' column names
  dplyr::rename_at(vars(everything()), ~ sub("max_mean", "max", .x)) %>% # remove '_mean' in 'maximum' column names
  dplyr::rename(n = oc_n) %>% # create 1 column to hold 'n'
  dplyr::relocate(n, .after = initDR) %>% 
  dplyr::select(-contains('_n'))
```



# 6. Emissions sources 
Sources of GHG emissions:  
c_fert_prod  
c_fert_appl  
c_pesticide     
c_field_energy_use  
  
Calculate differences due to each source.
```{r}
c_source <- c_gp_yld %>% 
  
  # find differences between BAU and MIT for each GHG source
  dplyr::group_by(region, soiltype, initDR, year) %>% 
  dplyr::mutate(
    # c_fert_prod
    c_fert_prod_diff = c_fert_prod - c_fert_prod[scenario == "BAU"],
    c_fert_prod_pc_change = (c_fert_prod_diff/c_fert_prod[scenario == "BAU"])*100,
    # c_fert_appl
    c_fert_appl_diff = c_fert_appl - c_fert_appl[scenario == "BAU"],
    c_fert_appl_pc_change = (c_fert_appl_diff/c_fert_appl[scenario == "BAU"])*100,
    # c_pesticide
    c_pesticide_diff = c_pesticide - c_pesticide[scenario == "BAU"],
    c_pesticide_pc_change = (c_pesticide_diff/c_pesticide[scenario == "BAU"])*100,
    # c_field_energy_use
    c_feu_diff = c_field_energy_use - c_field_energy_use[scenario == "BAU"],
    c_feu_pc_change = (c_feu_diff/c_field_energy_use[scenario == "BAU"])*100,
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(scenario == 'MIT') %>% 
   
  # get min and max values (this grouping will give range across years)----
  dplyr::group_by(region, soiltype, initDR) %>% 
  dplyr::mutate(
    # c_fert_prod
    c_fert_prod_diff_min = min(c_fert_prod_diff, na.rm = TRUE),
    c_fert_prod_diff_max = max(c_fert_prod_diff, na.rm = TRUE),
    # c_fert_appl
    c_fert_appl_diff_min = min(c_fert_appl_diff, na.rm = TRUE),
    c_fert_appl_diff_max = max(c_fert_appl_diff, na.rm = TRUE),
    # c_pesticide
    c_pesticide_diff_min = min(c_pesticide_diff, na.rm = TRUE),
    c_pesticide_diff_max = max(c_pesticide_diff, na.rm = TRUE),
    # c_field_energy_use
    c_feu_diff_min = min(c_feu_diff, na.rm = TRUE),
    c_feu_diff_max = max(c_feu_diff, na.rm = TRUE)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(ID:crop, c_fert_prod_diff:c_feu_diff_max) %>% 
  
  # calculate mean values across a rotation---- 
dplyr::group_by(scenario, region, soiltype, initDR) %>% 
  dplyr::summarise(across(c(
    # c_fert_prod
    c_fert_prod_diff, c_fert_prod_diff_min, c_fert_prod_diff_max,
    # c_fert_appl
    c_fert_appl_diff, c_fert_appl_diff_min, c_fert_appl_diff_max, 
    # c_pesticide
    c_pesticide_diff, c_pesticide_diff_min, c_pesticide_diff_max, 
    # c_field_energy_use
    c_feu_diff, c_feu_diff_min, c_feu_diff_max
    ),.fns = list(
      mean=~mean(.x, na.rm = TRUE),
      n=~nobs(.x)))) %>% 
  ungroup() %>% 
  
  # tidy up----
  dplyr::select(-contains(c('min_n', 'max_n'))) %>% 
  dplyr::rename_at(vars(everything()), ~ sub("min_mean", "min", .x)) %>% # remove '_mean' in 'minimum' column names
  dplyr::rename_at(vars(everything()), ~ sub("max_mean", "max", .x)) %>% # remove '_mean' in 'maximum' column names
  dplyr::rename(n = c_fert_prod_diff_n) %>% # create 1 column to hold 'n'
  dplyr::relocate(n, .after = initDR) %>% 
  dplyr::select(-contains('_n')) %>% 
  dplyr::ungroup()

c_source
```

Convert from wide to long for plotting.
```{r}
mean_df <- c_source %>%
  dplyr::select(region, soiltype, initDR, contains('mean')) %>% 
  dplyr::rename_at(vars(everything()), ~ sub("_diff_mean", "", .x)) %>% # remove '_diff_mean' from column names
  dplyr::rename_at(vars(everything()), ~ sub("c_", "", .x)) %>% # remove 'c_' from column names
  pivot_longer(
    #cols = ends_with("mean"),
    cols = fert_prod:feu,
    names_to = "source_of_emissions",
    #names_suffix = "_mean",
    values_to = "rotation_mean",
    values_drop_na = TRUE
  )

min_df <- c_source %>%
  dplyr::select(region, soiltype, initDR, contains('min')) %>% 
  dplyr::rename_at(vars(everything()), ~ sub("_diff_min", "", .x)) %>% # remove '_diff_mean' from column names
  dplyr::rename_at(vars(everything()), ~ sub("c_", "", .x)) %>% # remove 'c_' from column names
  pivot_longer(
    cols = fert_prod:feu,
    names_to = "source_of_emissions",
    values_to = "rotation_min",
    values_drop_na = TRUE
  )

max_df <- c_source %>%
  dplyr::select(region, soiltype, initDR, contains('max')) %>% 
  dplyr::rename_at(vars(everything()), ~ sub("_diff_max", "", .x)) %>% # remove '_diff_mean' from column names
  dplyr::rename_at(vars(everything()), ~ sub("c_", "", .x)) %>% # remove 'c_' from column names
  pivot_longer(
    cols = fert_prod:feu,
    names_to = "source_of_emissions",
    values_to = "rotation_max",
    values_drop_na = TRUE
  )

# Join data frames
sources <- dplyr::left_join(mean_df, min_df, by = c("region", "soiltype", "initDR", "source_of_emissions")) %>% 
  dplyr::left_join(., max_df, by = c("region", "soiltype", "initDR", "source_of_emissions")) %>% 
  #dplyr::mutate(across('source_of_emissions', str_replace, 'feu', 'field_energy_use'))
  dplyr::mutate(across('source_of_emissions', ~ str_replace(., 'feu', 'field_energy_use'))) %>% 
  dplyr::mutate(across('source_of_emissions', ~ str_replace(., 'fert_appl', 'fertiliser_application'))) %>% 
  dplyr::mutate(across('source_of_emissions', ~ str_replace(., 'fert_prod', 'fertiliser_production'))) 
  


# Tidy up
#rm(mean_df, min_df, max_df, c_source)
```


# 7. Plots
## 7.1. Direction of signs - what wins?
Difference in *gross profit: +ve value wins*  
Difference in gross profit (opp cost) = money lost by switching = difference in gross profit MIT-BAU.  
So if  
MIT = 10 <- better  
BAU = 5  
MIT - BAU = 5 = +ve value means MIT wins  
  
Difference in *carbon emissions: -ve value wins*  
Difference in carbon emissions MIT-BAU  
So if  
MIT = 5 <- better  
BAU = 8  
MIT - BAU = -3 = -ve value means MIT wins  
  
Difference in *calories: +ve value wins*  
Difference in kcal produced per m2 = MIT-BAU  
So if  
MIT = 6  <- better  
BAU = 3  
MIT - BAU = 3 = +ve value means MIT wins  
  
Difference in *land use: -ve value wins*    
Difference in amount of land used to produce 1000 kcal = MIT-BAU
So if  
MIT = 3  <- better  
BAU = 6  
MIT - BAU = -3 = -ve value means MIT wins
  
Could switch the sign on opportunity costs (would then be BAU-MIT) so that for both C emissions and gross profit, values below 0 mean MIT wins.  

  
  
## 7.2. Figure 1 - All outcomes panel plots
### 7.2.1. Specify panel labels
  
```{r, specify-panel-labels, echo=FALSE}
region.labs <- c("Northern England", "Central England", "Eastern England") # new
names(region.labs) <- c("north", "central", "east") # original

soiltype.labs <- c("Heavy soil", "Medium soil", "Light soil") # new
names(soiltype.labs) <- c("heavy", "medium", "light") # original
```
  
  

### 7.2.5. Combined plots

```{r}
# opp costs ----
a_oc_1col <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
               aes(initDR, oc_mean, colour = oc_mean < 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ region,
             labeller = labeller(soiltype = soiltype.labs, region = region.labs),
             scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = "Mean annual difference in gross profit (£/ha)") +
  geom_pointrange(aes(ymin = oc_min, ymax = oc_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        #axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        legend.position = "none",
        tag.text = element_text(size=22)
        ) 

# yield ----
b_cal_1col <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
               aes(initDR, calories_diff_mean, colour = calories_diff_mean < 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ region,
             labeller = labeller(soiltype = soiltype.labs, region = region.labs),
             scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = "Mean annual difference in calories produced (kcal/ha)") +
  geom_pointrange(aes(ymin = calories_diff_min, ymax = calories_diff_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        #axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        legend.position = "none"
        )

# GHG emissions ----
c_ghg_1col <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
               aes(initDR, c_emiss_diff_mean, colour = c_emiss_diff_mean > 0)) + 
    scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ region,
             labeller = labeller(soiltype = soiltype.labs, 
                                 region = region.labs)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = expression("Mean annual difference in GHG emissions (kgCO"["2"]~"eq/ha)"),
             scales="free_y"
             ) +
  geom_pointrange(aes(ymin = c_emiss_diff_min, ymax = c_emiss_diff_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        legend.position = "bottom"
        )

# combine ----
facetpanel_1col <- a_oc_1col + b_cal_1col + c_ghg_1col + plot_layout(ncol = 1) +
  plot_annotation(tag_levels = 'a', tag_prefix = '(',  tag_suffix = ')') & 
  theme(plot.tag = element_text(size = 22, face="bold"))
```





```{r, echo=FALSE, eval = FALSE}
# png("../figures/plot_all_outcomes_1col.png",
#     height = 30, width = 10, units = "in", res = 300)
# print(facetpanel_1col)
# dev.off()

pdf("../figures/plot_all_outcomes_1col.pdf", height = 30, width = 10)
print(facetpanel_1col)
dev.off()
```


## 7.3. Fig S?? 

```{r}

#library(ggbreak) 
figSx_yield <- ggplot(trdoff_rot, aes(initDR, yield_mean, colour = scenario, shape = scenario)) + 
  facet_grid(region ~ soiltype,
             labeller = labeller(region = region.labs, 
                                 soiltype = soiltype.labs)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  ylim(0, 10) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = "Mean annual crop yield (t/ha)") +
  # specify colours for 'scenario':
  scale_color_manual(name = "Scenario",
                     values = c("#00306F", # BAU, madison blue
                                "#00C2F9") # MIT, capri blue
                     ) + 
  scale_shape_manual(name = "Scenario",
                     values=c(15, # BAU, 15 = square
                              16) # MIT, 16 = circle
                     ) + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.title=element_text(size=24),
        legend.text=element_text(size=24)
        ) +
  geom_pointrange(aes(ymin = yield_min, ymax = yield_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9)

figSx_yield #+ scale_y_break(c(20, 60))
```

Save.
```{r}
png("../figures/Figure_Sx_yield-t-ha.png",
    height = 10, width = 10, units = "in", res = 300)
print(figSx_yield)
dev.off()
```


```{r}
E_L <- ggplot(trdoff_rot %>% dplyr::filter(region == "east" & soiltype == "light"), 
       aes(initDR, yield_mean, colour = scenario, shape = scenario)) + 
  #facet_grid(region ~ soiltype,
  #           labeller = labeller(region = region.labs, 
  #                               soiltype = soiltype.labs)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = "Mean annual wheat yield (t/ha)") +
  # specify colours for 'scenario':
  scale_color_manual(name = "Scenario",
                     values = c("#00306F", # BAU, madison blue
                                "#00C2F9") # MIT, capri blue
                     ) + 
  scale_shape_manual(name = "Scenario",
                     values=c(15, # BAU, 15 = square
                              16) # MIT, 16 = circle
                     ) + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.title=element_text(size=24),
        legend.text=element_text(size=24)
        ) +
  geom_pointrange(aes(ymin = yield_min, ymax = yield_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) + 
  #scale_y_break(c(20, 60))
  #scale_y_break(c(20, 60), space = 0.1, ticklabels = c(seq(0,80,20), 70), expand = c(0, 0))
  #scale_y_break(c(16, 67), scales = 0.5) # THIS WORKS OK ISH
  # this works better:
  scale_y_break(c(8, 15), ticklabels = c(seq(15,16,1), 16), scales = 0.3) + 
  scale_y_break(c(16, 67), ticklabels = c(seq(67,69,1), 70), scale=0.5) +
  theme(axis.text.x.top = element_blank(),
        axis.ticks.x.top = element_blank(),
        axis.line.x.top = element_blank())
  #scale_y_cut(breaks=c(8, 20), which=c(1, 3), scales=c(0.5, 3))

E_L
```




## 7.3. Figure S8 - Plot change in GHG emissions by source
### 7.3.1. Specify panel labels
  
```{r, specify-panel-labels, echo=FALSE}
denres.labs <- c("LD-LR", "LD-HR", "HD-HR") # new
names(denres.labs) <- c("LD-LR", "LD-HR", "HD-HR") # original

soiltype.labs <- c("Heavy soil", "Medium soil", "Light soil") # new
names(soiltype.labs) <- c("heavy", "medium", "light") # original
```

### 7.3.5. Combine plots
-ve value means MIT wins (green = MIT wins, orange = MIT loses) 
```{r}
# north ----
sources_n_1col <- ggplot(sources %>% dplyr::filter(region == "north"), 
               aes(source_of_emissions, rotation_mean, colour = rotation_mean > 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"), labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ initDR, 
             labeller = labeller(soiltype = soiltype.labs, initDR = denres.labs)#,
             #scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Source of GHG emissions", 
       y = expression("Difference in GHG emissions (kgCO"[2]~"eq / ha)")
       #y = expression(atop("Mean annual difference in",
      #                    paste("GHG emissions (kgCO"["2"]~"eq / ha)"))) 
       ) +
  geom_pointrange(aes(ymin = rotation_min, ymax = rotation_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        legend.position = "none",
        tag.text = element_text(size=22)
        ) 

# central ----
sources_c_1col <- ggplot(sources %>% dplyr::filter(region == "central"), 
               aes(source_of_emissions, rotation_mean, colour = rotation_mean > 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ initDR,
             labeller = labeller(soiltype = soiltype.labs, 
                                 initDR = denres.labs)
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Source of GHG emissions", 
       y = expression("Difference in GHG emissions (kgCO"[2]~"eq / ha)")
       #y = expression(atop("Mean annual difference in",
      #                    paste("GHG emissions (kgCO"["2"]~"eq / ha)"))) 
       ) +
  geom_pointrange(aes(ymin = rotation_min, ymax = rotation_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        legend.position = "none"
        )

# east ----
sources_e_1col <- ggplot(sources %>% dplyr::filter(region == "east"), 
               aes(source_of_emissions, rotation_mean, colour = rotation_mean > 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ initDR,
             labeller = labeller(soiltype = soiltype.labs, 
                                 initDR = denres.labs)
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Source of GHG emissions", 
       y = expression("Difference in GHG emissions (kgCO"[2]~"eq / ha)")
       #y = expression(atop("Mean annual difference in",
      #                    paste("GHG emissions (kgCO"["2"]~"eq / ha)"))) 
       ) +
  geom_pointrange(aes(ymin = rotation_min, ymax = rotation_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        legend.position = "bottom"
        )

# combine
sources_1col <- sources_n_1col + sources_c_1col + sources_e_1col + plot_layout(ncol = 1) +
  plot_annotation(
    #tag_levels = 'a', 
    tag_levels = list(c("(a) North", "(b) Central" , "(c) East")),
    tag_prefix = '',  
    tag_suffix = ''
    ) & 
  theme(plot.tag = element_text(size = 22, face="bold"))
```


```{r, echo=FALSE, eval = FALSE}
# png("../figures/plot_all_outcomes_1col.png",
#     height = 30, width = 10, units = "in", res = 300)
# print(facetpanel_1col)
# dev.off()

pdf("../figures/emissions_sources_1col.pdf", height = 30, width = 10)
print(sources_1col)
dev.off()
```

## 7.4. Density data violin plot

### 7.4.1. Using summarised data
```{r}
density_3dr <- readRDS("../data/density_data_summarised_3dr.rds") %>% 
  dplyr::filter(scenario != "CWW")

p <- ggplot(density_3dr %>% dplyr::filter(initDR3 == "LD-LR", year == 6), aes(factor(scenario), hvh_mean, fill = scenario))
p + geom_violin() + 
  geom_jitter(height = 0, width = 0.1)

```

### 7.4.2. Using raw simulation data
Re-do it with raw simulations. Sum high and very high density proportions (-> hvh) across 3 density-resistance levels.  
```{r}
# tidied simulated density data, all 37 imputations
res_tidy <- readRDS(file = "../data/simulation_results_tidy.rds") %>%  
  dplyr::filter(scenario != "CWW") %>% 
  dplyr::mutate( 
    scenario = factor(scenario, levels = c("BAU", "MIT")),
    initden = factor(initden, levels = c("VD", "HD", "MD", "LD")),
    initres = factor(initres),
    region = factor(region, levels = c("north", "central", "east")),
    soiltype = factor(soiltype, levels = c("heavy", "medium", "light")),
    initdenres = factor(initdenres, # specify order of levels
                        levels = c("VD-HR", "HD-HR", "MD-HR", 
                                   "MD-LR", "LD-HR", "LD-LR")),
    initcondit = factor(initcondit, levels = c(
      "n_h_VD-HR", "n_m_VD-HR", "n_l_VD-HR", 
      "c_h_VD-HR", "c_m_VD-HR", "c_l_VD-HR",
      "e_h_VD-HR", "e_m_VD-HR", "e_l_VD-HR",
      
      "n_h_HD-HR", "n_m_HD-HR", "n_l_HD-HR",
      "c_h_HD-HR", "c_m_HD-HR", "c_l_HD-HR",
      "e_h_HD-HR", "e_m_HD-HR", "e_l_HD-HR",
      
      "n_h_MD-HR", "n_m_MD-HR", "n_l_MD-HR",
      "c_h_MD-HR", "c_m_MD-HR", "c_l_MD-HR",
      "e_h_MD-HR", "e_m_MD-HR", "e_l_MD-HR",
      
      "n_h_MD-LR", "n_m_MD-LR", "n_l_MD-LR",
      "c_h_MD-LR", "c_m_MD-LR", "c_l_MD-LR",
      "e_h_MD-LR", "e_m_MD-LR", "e_l_MD-LR",
      
      "n_h_LD-HR", "n_m_LD-HR", "n_l_LD-HR",
      "c_h_LD-HR", "c_m_LD-HR", "c_l_LD-HR",
      "e_h_LD-HR", "e_m_LD-HR", "e_l_LD-HR",
      
      "n_h_LD-LR", "n_m_LD-LR", "n_l_LD-LR",
      "c_h_LD-LR", "c_m_LD-LR", "c_l_LD-LR",
      "e_h_LD-LR", "e_m_LD-LR", "e_l_LD-LR")),
    strategy = factor(strategy)
  ) %>%
  ungroup() %>%
  # collapse init den res to three levels
  dplyr::mutate(
    initDR3 = forcats::fct_recode(initdenres, # collapse categories
                                  "LD-LR" = "MD-LR", # new, old
                                  "LD-HR" = "MD-HR", 
                                  "HD-HR" = "VD-HR"),
    initDR3 = factor(initDR3, levels = c("LD-LR", "LD-HR", "HD-HR"))
                ) %>% 
  ungroup() %>% 
  # sum the proportions of high and very high density black-grass:
  rowwise() %>% 
  dplyr::mutate(hvh = sum(c_across(h:v), na.rm = T)) %>%  # new category: h & vh
  ungroup() %>% 
  dplyr::rename(mds = mds.ms)
```


Plot LD-LR
```{r}
p_LDLR <- ggplot(res_tidy %>% filter(initDR3 == "LD-LR", year == 6), 
             aes(x = factor(scenario), y = hvh, fill = scenario)) +  
  geom_violin(alpha = 0.5) +  # Semi-transparent violin plots
  geom_jitter(width = 0.1, height = 0, color = "lightgrey", size = 0.5, alpha = 0.5) +  # Tiny, light grey, transparent points
  labs(x = "Scenario", y = "Proportion of field with high or very \nhigh density black-grass at year 6") +  # Custom axis titles
  scale_fill_manual(values = c("MIT" = "#40B0A6", "BAU" = "#E1BE6A")) +  # Custom colors for MIT and BAU
  theme_minimal() +  # Base minimal theme for a clean look
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(color = "black"),  # Black axis lines for both x and y
    axis.ticks = element_line(color = "black"),  # Black axis ticks
    axis.title=element_text(size=24),
    axis.text.x=element_text(size=rel(2.2)),
    axis.text.y=element_text(size=rel(2.2)),
    plot.title = element_text(size = 22, face = "bold"),
    #legend.title=element_text(size=22),
    #legend.text=element_text(size=20),
    legend.position = "none"  # Remove legend
  ) + 
  ggtitle("LD-LR") 

p_LDLR
```


Plot LD-HR
```{r}
p_LDHR <- ggplot(res_tidy %>% filter(initDR3 == "LD-HR", year == 6), 
             aes(x = factor(scenario), y = hvh, fill = scenario)) +  
  geom_violin(alpha = 0.5) +  # Semi-transparent violin plots
  geom_jitter(width = 0.1, height = 0, color = "lightgrey", size = 0.5, alpha = 0.5) +  # Tiny, light grey, transparent points
  labs(x = "Scenario", y = "Proportion of field with high or \nvery high density black-grass at year 6") +  # Custom axis titles
  scale_fill_manual(values = c("MIT" = "#40B0A6", "BAU" = "#E1BE6A")) +  # Custom colors for MIT and BAU
  theme_minimal() +  # Base minimal theme for a clean look
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(color = "black"),  # Black axis lines for both x and y
    axis.ticks = element_line(color = "black"),  # Black axis ticks
    axis.title=element_text(size=24),
    axis.title.y = element_blank(),
    axis.text.x=element_text(size=rel(2.2)),
    axis.text.y=element_text(size=rel(2.2)),
    plot.title = element_text(size = 22, face = "bold"),
    legend.position = "none"  # Remove legend
  ) + 
  ggtitle("LD-HR") 
  
p_LDHR
```


Plot HD-HR
```{r}
p_HDHR <- ggplot(res_tidy %>% filter(initDR3 == "HD-HR", year == 6), 
             aes(x = factor(scenario), y = hvh, fill = scenario)) +  
  geom_violin(alpha = 0.5) +  # Semi-transparent violin plots
  geom_jitter(width = 0.1, height = 0, color = "lightgrey", size = 0.5, alpha = 0.5) +  # Tiny, light grey, transparent points
  labs(x = "Scenario", y = "Proportion of field with high or \nvery high density black-grass at year 6") +  # Custom axis titles
  scale_fill_manual(values = c("MIT" = "#40B0A6", "BAU" = "#E1BE6A")) +  # Custom colors for MIT and BAU
  theme_minimal() +  # Base minimal theme for a clean look
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    panel.grid = element_blank(),  # Remove gridlines
    axis.line = element_line(color = "black"),  # Black axis lines for both x and y
    axis.ticks = element_line(color = "black"),  # Black axis ticks
    axis.title=element_text(size=24),
    axis.title.y = element_blank(),
    axis.text.x=element_text(size=rel(2.2)),
    axis.text.y=element_text(size=rel(2.2)),
    plot.title = element_text(size = 22, face = "bold"),
    legend.position = "none"  # Remove legend
  ) + 
  ggtitle("HD-HR") 
  
p_HDHR
```


Create panel plot.
```{r}
bg_density_yr6 <- (p_LDLR + p_LDHR + p_HDHR) + plot_layout(axis_titles = "collect") 
bg_density_yr6
```




Save panel plot.
```{r}
png("../figures/bg_density_yr6.png",
    height = 6, width = 18, units = "in", res = 300)
print(bg_density_yr6)
dev.off()
```


How many points are in each violin plot?
```{r}
nrow(res_tidy %>% filter(scenario == "MIT", initDR3 == "HD-HR", year == 6))
```

```{r}
n_distinct(res_tidy$initcondit) # i.e., e_l_HD-HR
```

```{r}
unique(res_tidy$initcondit)
```


```{r}
unique(res_tidy$initdenres) # i.e., LD-HR or VD-HR etc.
```



## 7.5. Diversification plot
```{r}
# load data
strat <- read_xlsx("../data/strategy design summary.xlsx",2) %>% 
  dplyr::mutate(
    scenario = as.factor(scenario),
    design_element = as.factor(design_element),
    initdenres = as.factor(initdenres),
    region = as.factor(region),
    soil_type = as.factor(soil_type)
  ) 
 
glimpse(strat)
```


```{r}
unique(strat$design_element)
```

Calculate total volume of glyphosate applied in a rotation.
```{r}
strat2 <- strat %>% 
  # calculate total glyphosate volume in a rotation
  filter(design_element == "gly_vol_yr") %>%
  dplyr::group_by(scenario, design_element, initdenres, region, soil_type) %>% 
  summarise(value = sum(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  dplyr::mutate(
    # rename design element so it is now gly_rot_vol
    design_element = forcats::fct_recode(design_element, "gly_rot_vol" = "gly_vol_yr"),
    # add 'year' and 'scaled_value' columns so can bind to original data frame
    year = 'NA'
    ) %>% 
  dplyr::relocate(year, .before = value)

colnames(strat2)
# join the two dataframes
strat <- rbind(strat, strat2)

strat <- strat %>% 
  # Within each design element, scale values between 0 and 1:
  group_by(design_element) %>%
  dplyr::mutate(scaled_value = (value - min(value)) / (max(value) - min(value))) %>%
  ungroup() 

rm(strat2)

glimpse(strat)
```


Plot several design elements together.
```{r}
strat %>%
  dplyr::mutate(initdenres = fct_recode(initdenres, "BAU" = "NA")) %>% 
  dplyr::mutate(initdenres = factor(initdenres, levels=c("BAU", "LD-LR", "LD-HR", "HD-HR"))) %>%
  dplyr::mutate(ID = paste(scenario, initdenres)) %>% 
  dplyr:: group_by(ID) %>% 
  dplyr::filter(design_element %in% c("tillage_depth_yr", "herb_applic_n_yr", "selherb_vol_yr", "gly_vol_yr" )) %>% #"aut_gly_n_yr", 
  dplyr::mutate(design_element = factor(design_element, levels=c("tillage_depth_yr", "herb_applic_n_yr", "selherb_vol_yr",  "gly_vol_yr"))) %>% #"aut_gly_n_yr",
  ggplot(aes(fill=design_element, y=scaled_value, x=initdenres)) +
    #geom_point(color="lightgrey", size=1, width=0.1, position=position_jitterdodge() , alpha=0.4) +
    geom_boxplot(position=position_dodge2(preserve = "total"), alpha=0.5, outlier.colour="grey") + #, outlier.colour="transparent"
    scale_fill_viridis(discrete=T, name="",
                       labels = c("Tillage depth", 
                                  "Herbicide application frequency", 
                                  "Selective herbicide volume", 
                                  #"Autumn glyphosate application frequency",  
                                  "Glyphosate volume")
                       ) +
    #geom_text(data=strat, aes(label=paste0("n: ",n), y=median-2), position=position_dodge(1), hjust=0.5) +
    #theme_ipsum()  +
    theme_classic() +
    xlab("") +
    ylab("Scaled value") #+
    #ylim(0,1)
```

```{r}
unique(strat$design_element)
```

Another one for rotation values.
```{r}


strat3 %>%
  dplyr::mutate(initdenres = fct_recode(initdenres, "BAU" = "NA")) %>% 
  dplyr::mutate(initdenres = factor(initdenres, levels=c("BAU", "LD-LR", "LD-HR", "HD-HR"))) %>%
  dplyr::mutate(ID = paste(scenario, initdenres)) %>% 
  dplyr:: group_by(ID) %>% 
  dplyr::filter(design_element %in% c("diff_act_tot_n_rot", "diff_crops_tot_n_rot", "spring_crops_tot_n_rot", "ssb_tot_n_rot", "gly_rot_vol" )) %>%  
  #dplyr::mutate(design_element = factor(design_element, levels=c("tillage_depth_yr", "herb_applic_n_yr", "selherb_vol_yr",  "gly_vol_yr"))) %>% 
  ggplot(aes(fill=design_element, y=value, x=initdenres)) +
    #geom_point(color="lightgrey", size=1, width=0.1, position=position_jitterdodge() , alpha=0.4) +
    geom_boxplot(position=position_dodge2(preserve = "total"), alpha=0.5, outlier.colour="grey") + #, outlier.colour="transparent"
    scale_fill_viridis(discrete=T, name="",
                       labels = c("Number of different actives", 
                                  "Number of different crops", 
                                  "Number of spring crops", 
                                  "Number of stale seed beds",  
                                  "Glyphosate total volume")
                       ) +
    #geom_text(data=strat, aes(label=paste0("n: ",n), y=median-2), position=position_dodge(1), hjust=0.5) +
    #theme_ipsum()  +
    theme_classic() +
    xlab("") +
    ylab("Value") #+
    #ylim(0,1)
```




Plot a lollipop for rotation values?
```{r}
strat %>%
  dplyr::mutate(initdenres = fct_recode(initdenres, "BAU" = "NA")) %>%
  dplyr::mutate(initdenres = factor(initdenres, levels = c("BAU", "LD-LR", "LD-HR", "HD-HR"))) %>%
  dplyr::mutate(ID = paste(scenario, initdenres)) %>%
  dplyr::filter(design_element %in% c("tillage_depth_yr", "herb_applic_n_yr", "aut_gly_n_yr", 
                                     "selherb_vol_yr", "gly_vol_yr", "diff_crops_tot_n_rot")) %>%
  dplyr::group_by(scenario, initdenres, design_element) %>%
  dplyr::summarize(mean_value = mean(scaled_value, na.rm = TRUE)) %>%  # Calculate mean for each group
  ungroup() %>% 
  group_by(design_element) %>%
  dplyr::mutate(scaled_mean_value = (mean_value - min(mean_value)) / (max(mean_value) - min(mean_value))) %>%
  ggplot(aes(x = initdenres, y = scaled_mean_value, color = design_element, group = design_element)) +
  geom_segment(aes(xend = initdenres, yend = 0), size = 1, position = position_dodge(width = 0.8)) +  # Adjust line position to avoid overlap
  geom_point(size = 4, aes(fill = design_element), shape = 21, stroke = 1.5, position = position_dodge(width = 0.8)) +  # Apply dodge to points too
  scale_color_viridis(discrete = TRUE, name = "", 
                      labels = c("Tillage depth", "Herbicide application frequency", 
                                 "Autumn glyphosate application frequency", 
                                 "Selective herbicide volume", "Glyphosate volume", 
                                 "Number of different crops")) +
  scale_fill_viridis(discrete = TRUE, guide = "none") +  # No fill color in the legend
  theme_classic() +
  xlab("") +
  ylab("Mean Value") +
  ylim(0, 1) +
  theme(legend.title = element_blank(), 
        legend.text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, hjust = 1))



ggplot(data, aes(x = Category, y = Value, color = Colour, group = SubCategory)) +
  geom_point(size = 5) + # Points
  geom_segment(aes(xend = Category, yend = 0), size = 1) + # Lines
  facet_wrap(~ Category, scales = "free_y", ncol = 1) + # Facet the categories
  theme_minimal() + # Clean theme
  labs(title = "Lollipop Plot with Categories and Sub-categories",
       x = "Category", y = "Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability

strat %>%
  dplyr::mutate(initdenres = fct_recode(initdenres, "BAU" = "NA")) %>%
  dplyr::mutate(initdenres = factor(initdenres, levels = c("BAU", "LD-LR", "LD-HR", "HD-HR"))) %>%
  dplyr::mutate(ID = paste(scenario, initdenres)) %>%
  dplyr::filter(design_element %in% c("tillage_depth_yr", "herb_applic_n_yr", "aut_gly_n_yr", 
                                     "selherb_vol_yr", "gly_vol_yr", "diff_crops_tot_n_rot")) %>%
  dplyr::group_by(scenario, initdenres, design_element) %>%
  dplyr::summarize(mean_value = mean(value, na.rm = TRUE)) %>%  # Calculate mean for each group
  dplyr::mutate()
  
  ggplot(aes(x = initdenres, y = mean_value, color = design_element, group = design_element)) +
  geom_segment(aes(xend = initdenres, yend = 0), size = 1, position = position_dodge(width = 0.8)) +  # Adjust line position to avoid overlap
  geom_point(size = 4, aes(fill = design_element), shape = 21, stroke = 1.5, position = position_dodge(width = 0.8)) +  # Apply dodge to points too
  scale_color_viridis(discrete = TRUE, name = "", 
                      labels = c("Tillage depth", "Herbicide application frequency", 
                                 "Autumn glyphosate application frequency", 
                                 "Selective herbicide volume", "Glyphosate volume", 
                                 "Number of different crops")) +
  scale_fill_viridis(discrete = TRUE, guide = "none") +  # No fill color in the legend
  theme_classic() +
  xlab("") +
  ylab("Mean Value") +
  ylim(0, 1) +
  theme(legend.title = element_blank(), 
        legend.text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, hjust = 1))


```



Re-do with design element on x axis and bars showing initdenres.
```{r}
strat %>%
  dplyr::mutate(initdenres = fct_recode(initdenres, "BAU" = "NA")) %>% 
  dplyr::mutate(initdenres = factor(initdenres, levels=c("BAU", "LD-LR", "LD-HR", "HD-HR"))) %>%
  dplyr::mutate(ID = paste(scenario, initdenres)) %>% 
  dplyr:: group_by(ID) %>% 
  dplyr::filter(design_element %in% c("tillage_depth_yr", "herb_applic_n_yr", "aut_gly_n_yr", "selherb_vol_yr", "gly_vol_yr", "diff_crops_tot_n_rot" )) %>% 
  ggplot(aes(fill=initdenres, y=scaled_value, x=design_element)) +
    geom_boxplot(position=position_dodge2(preserve = "total"), alpha=0.5, outlier.colour="transparent") +
    #geom_point(color="grey", size=1, width=0.1, position=position_jitterdodge() , alpha=0.4) +
    scale_fill_viridis(discrete=T, name="") +
    #geom_text(data=strat, aes(label=paste0("n: ",n), y=median-2), position=position_dodge(1), hjust=0.5) +
    #theme_ipsum()  +
    theme_classic() +
    xlab("") +
    ylab("Scaled value") +
    ylim(0,1) +
  scale_x_discrete(labels = c(
    "aut_gly_n_yr" = "# autumn glyphosate\napp",
    "diff_crops_tot_n_rot" = "# diff\ncrops rot",
    "gly_vol_yr" = "glyphosate volume\nper yr",
    "herb_applic_n_yr" = "# herb app\nper yr",
    "selherb_vol_yr" = "vol selherb\nper yr",
    "tillage_depth_yr" = "tillage depth"
  ))
```

# Cite packages
```{r}
cite_packages(citation.style = "nature", out.dir = ".")
```
See here https://pakillo.github.io/grateful/index.html for instructions on using `grateful`.



# Superseded
## 3.3. NOT USED - Calculate % change from individual years
Here, I calculate the difference between BAU and MIT for each year, then calculate the % change for each year, then take the average % change.
```{r}
# pc_change <- c_gp_yld %>%
#   
#   # 1. Calculate percentage change for each year of each scenario
#   dplyr::group_by(region, soiltype, initDR, year) %>% 
#   dplyr::mutate(
#     
#     # gross profit % change
#     gp_pc_change = (oc/gp[scenario == "BAU"])*100, # opportunity cost is already the difference in gross profit 
#     
#     # calories % change
#     cal_pc_change = (calories_diff/calories[scenario == "BAU"])*100,    
#     
#     # land use % change
#     lu_pc_change = (land_use_diff/land_use[scenario == "BAU"])*100,  
#     
#     # carbon emissions % change
#     c_pc_change = (c_emiss_diff/c_total[scenario == "BAU"])*100
#     
#   ) %>% 
#   ungroup() %>% 
#   
#   # 2. calculate min and max values (this grouping will give range across years)
#   dplyr::group_by(scenario, region, soiltype, initDR, year) %>% 
#   dplyr::mutate(
#     
#     # gross profit difference min & max
#     gp_diff_min = min(oc, na.rm = TRUE),
#     gp_diff_max = max(oc, na.rm = TRUE),
#     
#     # calories difference min & max
#     calories_diff_min = min(calories_diff, na.rm = TRUE),
#     calories_diff_max = max(calories_diff, na.rm = TRUE),
#     
#     # land use difference min & max
#     land_use_diff_min = min(land_use_diff, na.rm = TRUE),
#     land_use_diff_max = max(land_use_diff, na.rm = TRUE),
#     
#     # GHG difference min & max
#     c_emiss_diff_min = min(c_emiss_diff, na.rm = TRUE),
#     c_emiss_diff_max = max(c_emiss_diff, na.rm = TRUE)
#     
#   ) %>% 
#   ungroup() %>% 
#   
#   # 3. Calculate mean for each level of initDR
#   dplyr::filter(scenario == "MIT") %>% 
#   dplyr::group_by(initDR) %>% 
#   dplyr::mutate(
#     # gross profit mean, min, max % change
#     gp_pcchange_mean = mean(gp_pc_change),
#     gp_pcchange_min = min(gp_pc_change),
#     gp_pcchange_max = max(gp_pc_change),
#     
#     # calories mean min, max % change
#     cal_pcchange_mean = mean(cal_pc_change),
#     cal_pcchange_min = min(cal_pc_change),
#     cal_pcchange_max = max(cal_pc_change),
#     
#     # land use mean min, max % change
#     lu_pcchange_mean = mean(lu_pc_change),
#     lu_pcchange_min = min(lu_pc_change),
#     lu_pcchange_max = max(lu_pc_change),
#     
#     # carbon emissions mean min, max % change
#     c_pcchange_mean = mean(c_pc_change),
#     c_pcchange_min = min(c_pc_change),
#     c_pcchange_max = max(c_pc_change)
#     
#   ) %>% 
#   
#   #Select just useful columns and filter out duplicates
#   dplyr::select(initDR, contains("pcchange")) %>% 
#   dplyr::distinct(initDR, .keep_all = TRUE) %>% 
#   arrange(initDR)
# 
# pc_change
```

## 3.4. Save output as .csv
```{r}
#write.csv(pc_change, "../output/summarytable_from_indiv_years.csv", row.names = FALSE)
```

## 7.alt Mean difference in C emissions per initDR
This gives min and max values for any single year of a management strategy. *This is NOT what I use in the paper.*  
```{r}
c_gp_yld %>%
  dplyr::filter(scenario == 'MIT') %>% 
  dplyr::group_by(initDR) %>% 
  dplyr::summarize(
    #gp_mean = mean(gp),
    #cal_mean = mean(calories),
    c_emiss_meandiff = mean(c_emiss_diff),
    c_emiss_min = min(c_emiss_diff),
    c_emiss_max = max(c_emiss_diff)
  ) %>% 
  ungroup() 
``` 
### 7.2.2. Difference in gross profit
+ve value means MIT wins (green = MIT wins, orange = MIT loses)  
```{r, echo=FALSE}
facet_oc <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
               aes(initDR, oc_mean, colour = oc_mean < 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      #name = "MIT wins?",
                      labels = c("MIT wins", "MIT loses")) +
                      #labels = c("Yes", "No")) +
  facet_grid(soiltype ~ region,
             labeller = labeller(soiltype = soiltype.labs, 
                                 region = region.labs)#,
             #scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = "Mean annual difference in gross profit (£/ha)" 
       ) +
  geom_pointrange(aes(ymin = oc_min, ymax = oc_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank()
        )

facet_oc
```
  
  

```{r, echo=FALSE, eval = FALSE}
# png("../figures/plot_oc.png",
#     height = 15, width = 15, units = "in", res = 300)
# print(facet_oc)
# dev.off()

pdf("../figures/plot_oc.pdf", height = 15, width = 15)
print(facet_oc)
dev.off()
```



### 7.2.3. Difference in yield (calories)
+ve value means MIT wins
```{r, echo=FALSE}
facet_cal <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
               aes(initDR, calories_diff_mean, colour = calories_diff_mean < 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ region,
             labeller = labeller(soiltype = soiltype.labs, 
                                 region = region.labs)#,
             #scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = "Mean annual difference in calories produced (kcal/ha)" 
       ) +
  geom_pointrange(aes(ymin = calories_diff_min, ymax = calories_diff_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank()
        )

facet_cal
```
  
  

```{r, echo=FALSE, eval = FALSE}
# png("../figures/plot_cal.png",
#     height = 15, width = 15, units = "in", res = 300)
# print(facet_cal)
# dev.off()

pdf("../figures/plot_cal.pdf", height = 15, width = 15)
print(facet_cal)
dev.off()
```



### 7.2.4. Difference in GHG emissions
-ve value means MIT wins
```{r, echo=FALSE}
facet_c <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
               aes(initDR, c_emiss_diff_mean, colour = c_emiss_diff_mean > 0)) + 
    scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ region,
             labeller = labeller(soiltype = soiltype.labs, 
                                 region = region.labs)#,
             #scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Initial density and resistance of Black-grass population", 
       y = expression("Mean annual difference in GHG emissions (kgCO"["2"]~"eq/ha)") 
       ) +
  geom_pointrange(aes(ymin = c_emiss_diff_min, ymax = c_emiss_diff_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank()
        )
facet_c
```
  
  

```{r, echo=FALSE, eval = FALSE}
# png("../figures/plot_GHG.png",
#     height = 15, width = 15, units = "in", res = 300)
# print(facet_c)
# dev.off()

pdf("../figures/plot_GHG.pdf", height = 15, width = 15)
print(facet_c)
dev.off()
```




### 7.2.5. alt - Square
```{r}
# # opp costs ----
# a_oc <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
#                aes(initDR, oc_mean, colour = oc_mean < 0)) + 
#   scale_colour_manual(values = c("#009E73", "#D55E00"),
#                       labels = c("MIT wins", "MIT loses")) +
#   facet_grid(soiltype ~ region,
#              labeller = labeller(soiltype = soiltype.labs, region = region.labs)) +
#   scale_x_discrete(guide = guide_axis(angle = 45)) +
#   labs(x = "Initial density and resistance of Black-grass population", 
#        y = "Mean annual difference in gross profit (£/ha)") +
#   geom_pointrange(aes(ymin = oc_min, ymax = oc_max), 
#                   position=position_dodge(width=0.3),
#                   size = 0.9) +
#   geom_hline(yintercept=0, linetype="dashed", color = "black") +
#   theme_bw() + 
#   theme(strip.placement = 'outside',
#         strip.text.x = element_text(size = 22),
#         strip.text.y = element_text(size = 22),
#         axis.title=element_text(size=24),
#         axis.text.x=element_text(size=rel(2.2)),
#         axis.text.y=element_text(size=rel(2.2)),
#         legend.text=element_text(size=22),
#         legend.title = element_blank(),
#         legend.position = "none"
#         ) 
# 
# # yield ----
# b_cal <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
#                aes(initDR, calories_diff_mean, colour = calories_diff_mean < 0)) + 
#   scale_colour_manual(values = c("#009E73", "#D55E00"),
#                       labels = c("MIT wins", "MIT loses")) +
#   facet_grid(soiltype ~ region,
#              labeller = labeller(soiltype = soiltype.labs, region = region.labs)) +
#   scale_x_discrete(guide = guide_axis(angle = 45)) +
#   labs(x = "Initial density and resistance of Black-grass population", 
#        y = "Mean annual difference in calories produced (kcal/ha)") +
#   geom_pointrange(aes(ymin = calories_diff_min, ymax = calories_diff_max), 
#                   position=position_dodge(width=0.3),
#                   size = 0.9) +
#   geom_hline(yintercept=0, linetype="dashed", color = "black") +
#   theme_bw() + 
#   theme(strip.placement = 'outside',
#         strip.text.x = element_text(size = 22),
#         strip.text.y = element_text(size = 22),
#         axis.title=element_text(size=24),
#         axis.text.x=element_text(size=rel(2.2)),
#         axis.text.y=element_text(size=rel(2.2)),
#         legend.text=element_text(size=22),
#         legend.title = element_blank(),
#         legend.position = "none"
#         )
# 
# # GHG emissions ----
# c_ghg <- ggplot(trdoff_rot %>% dplyr::filter(scenario == "MIT"), 
#                aes(initDR, c_emiss_diff_mean, colour = c_emiss_diff_mean > 0)) + 
#     scale_colour_manual(values = c("#009E73", "#D55E00"),
#                       labels = c("MIT wins", "MIT loses")) +
#   facet_grid(soiltype ~ region,
#              labeller = labeller(soiltype = soiltype.labs, 
#                                  region = region.labs)) +
#   scale_x_discrete(guide = guide_axis(angle = 45)) +
#   labs(x = "Initial density and resistance of Black-grass population", 
#        y = expression("Mean annual difference in GHG emissions (kgCO"["2"]~"eq/ha)")) +
#   geom_pointrange(aes(ymin = c_emiss_diff_min, ymax = c_emiss_diff_max), 
#                   position=position_dodge(width=0.3),
#                   size = 0.9) +
#   geom_hline(yintercept=0, linetype="dashed", color = "black") +
#   theme_bw() + 
#   theme(strip.placement = 'outside',
#         strip.text.x = element_text(size = 22),
#         strip.text.y = element_text(size = 22),
#         axis.title=element_text(size=24),
#         axis.text.x=element_text(size=rel(2.2)),
#         axis.text.y=element_text(size=rel(2.2)),
#         legend.text=element_text(size=22),
#         legend.title = element_blank(),
#         legend.position = "none"
#         ) 
# 
# # get legend
# g_legend<-function(a.gplot){
#     tmp <- ggplot_gtable(ggplot_build(a.gplot))
#     leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
#     legend <- tmp$grobs[[leg]]
#     legend
# }
# 
# legend <- g_legend(facet_oc)
# 
# # combine
# facetpanel <- a_oc + b_cal + c_ghg + legend + plot_layout(ncol = 2)
```



```{r, echo=FALSE, eval = FALSE}
# png("../figures/plot_all_outcomes_square.png",
#     height = 20, width = 20, units = "in", res = 300)
# print(facetpanel)
# dev.off()

# pdf("../figures/plot_all_outcomes_square.pdf", height = 20, width = 20)
# print(facetpanel)
# dev.off()
```


### 7.3.2. Plot north region 
-ve value means MIT wins (green = MIT wins, orange = MIT loses)  
```{r, echo=FALSE}
eissions_sources_north <- ggplot(sources %>% dplyr::filter(region == "north"), 
               aes(source_of_emissions, rotation_mean, colour = rotation_mean > 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      #name = "MIT wins?",
                      labels = c("MIT wins", "MIT loses")) +
                      #labels = c("Yes", "No")) +
  facet_grid(soiltype ~ initDR,
             labeller = labeller(soiltype = soiltype.labs, 
                                 initDR = denres.labs)#,
             #scales="free_y"
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Source of GHG emissions", 
       y = expression("Mean annual difference in GHG emissions (kgCO"["2"]~"eq / ha)") 
       ) +
  geom_pointrange(aes(ymin = rotation_min, ymax = rotation_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank()
        )

eissions_sources_north
```
 


### 7.3.3. Plot central region 
-ve value means MIT wins (green = MIT wins, orange = MIT loses)  
```{r, echo=FALSE}
eissions_sources_central <- ggplot(sources %>% dplyr::filter(region == "central"), 
               aes(source_of_emissions, rotation_mean, colour = rotation_mean > 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ initDR,
             labeller = labeller(soiltype = soiltype.labs, 
                                 initDR = denres.labs)
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Source of GHG emissions", 
       y = expression("Mean annual difference in GHG emissions (kgCO"["2"]~"eq / ha)") 
       ) +
  geom_pointrange(aes(ymin = rotation_min, ymax = rotation_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank()
        )

eissions_sources_central
```
 


### 7.3.4. Plot east region 
-ve value means MIT wins (green = MIT wins, orange = MIT loses)  
```{r, echo=FALSE}
eissions_sources_east <- ggplot(sources %>% dplyr::filter(region == "east"), 
               aes(source_of_emissions, rotation_mean, colour = rotation_mean > 0)) + 
  scale_colour_manual(values = c("#009E73", "#D55E00"),
                      labels = c("MIT wins", "MIT loses")) +
  facet_grid(soiltype ~ initDR,
             labeller = labeller(soiltype = soiltype.labs, 
                                 initDR = denres.labs)
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "Source of GHG emissions", 
       y = expression("Mean annual difference in GHG emissions (kgCO"["2"]~"eq / ha)") 
       ) +
  geom_pointrange(aes(ymin = rotation_min, ymax = rotation_max), 
                  position=position_dodge(width=0.3),
                  size = 0.9) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  theme_bw() + 
  theme(strip.placement = 'outside',
        strip.text.x = element_text(size = 22),
        strip.text.y = element_text(size = 22),
        axis.title=element_text(size=24),
        axis.text.x=element_text(size=rel(2.2)),
        axis.text.y=element_text(size=rel(2.2)),
        legend.text=element_text(size=22),
        legend.title = element_blank()
        )

eissions_sources_east
```
  
