---
title: "Summarise & visualise results"
author: "Alexa Varah"
date: "2023-06-05"
output: html_document
---


# 1. Intro

## Sources of emissions in CFT

**Pesticide emissions**  
Pesticide emissions = number of pesticide applications * embodied emissions in pesticides (kg COs-eq / dose / ha) * crop yield / ha  
This doesn't account for the fuel used when applying pesticides, only the emissions caused by pesticide production.  

**Direct and indirect field N2O**  
Direct and indirect field N2O = Background direct and indirect N2O + Fertiliser induced field emissions (FIE)  

Background direct and indirect N2O = N2O + NO (they don't include NH3 volatilisation + leaching in this equation)
Fertiliser induced field emissions = N to N2O conversion factor * (N2O + NO + NH3 volatilisation + leaching)
NH3 volatilisation = volatilisation emissions factor of 0.01 * 
                      sum of NH3 volatilisation from each fertiliser application
NH3 volatilisation from a fertiliser application = amount of N applied * 
      EXP(sum of NH3 volatilisation factors for climate, crop type, soil texture, SOM, soil CEC, drainage, pH + 
              NH3 from fertiliser production emissions + 
                NH3 from fertiliser application)
              
**Field energy use**  
This contains emissions from diesel used in tillage, herb spraying, fert spreading/spraying, other field ops (harvest etc - we haven't input data on these other field ops).


# 2. Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 3. Load packages and data
## 3.1. Load packages
```{r, include=FALSE}
rm(list = ls(all = TRUE)) # make sure envt is clear
library(tidyverse)
library(viridis)
```

## 3.2. Load functions
```{r load-functions, include=FALSE}
# function to count complete observations
nobs <- function(x) length(x[!is.na(x)])
```

## 3.3. Load & tidy data
```{r, include=FALSE}
c_emis <- read.csv("../output/carbon_results.csv", header = T) %>% 
  dplyr::rename(ID = Field.ID,
                fert_prod = "Fertiliser.production..kgC02eq.ha.",
                fert_appl = "Field.N20.from.fertiliser..kgC02eq.ha.",
                pesticide = "Pesticide..kgC02eq.ha.",
                c_stock_change = "C.stock.changes..kgC02eq.ha.",
                field_energy_use = "Field.energy.use..kgC02eq.ha.",
                total = "Total..kgC02eq.ha.") %>%
  dplyr::mutate(to_split = ID) %>% 
  separate(to_split, 
                  into = c("scenario", "region", "soiltype",
                           "initDR", "year"),sep = "_") %>% 
  dplyr::mutate(
    scenario = factor(scenario, levels = c("BAU", "MIT", "CWW")),
    region = dplyr::recode(region, # rename levels of 'region'
                           "n" = "north",
                           "c" = "central",
                           "e" = "east"),
    region = factor(region, levels = c("north", "central", "east")),
    soiltype = dplyr::recode(soiltype,  # rename levels of 'soiltype'
                              "h" = "heavy" ,
                              "m" = "medium",
                              "l" = "light"),
    soiltype = factor(soiltype, levels = c("heavy", "medium","light")),
    initDR = factor(initDR, levels = c("LD-LR", "LD-HR", "HD-HR"))
    )
```

Save tidy results.
```{r}
write.csv(c_emis, "../output/carbon_results_tidy.csv", row.names = FALSE)
```


